---
title: "Sankey Diagram for Stroke Risk Factors in Blacks"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Sankey Diagram for Stroke Risk Factors in Blacks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 9,
  fig.height = 5
)
library(strokeSankey)
```

# Overview

This vignette demonstrates how the `strokeSankey` package recreates the Sankey-style diagram showing the contribution of major stroke risk factors among Black populations over time.

The data represent Population Attributable Risk (PAR) estimates for several risk factors across five years: 1990, 1995, 2000, 2005, and 2010.

The Sankey diagram shows how the contribution of each risk factor changes across time.

# Dataset

The dataset included in the package is `stroke_par_blacks`.

```{r}
data(stroke_par_blacks)
stroke_par_blacks
```

Each row corresponds to a stroke risk factor and the columns represent PAR estimates for each year.

# Creating the Sankey Diagram

The function `plot_sankey_stroke()` generates the Sankey diagram.

```{r}
plot_sankey_stroke(stroke_par_blacks)
```

The visualization is constructed using `ggplot2` by combining stacked rectangles for each year and flowing ribbons connecting the same risk factor across years.

# Controlling Whitespace

The function allows control over the whitespace between visual elements.

- `vertical_gap` controls the vertical spacing between stacked ribbons
- `horizontal_gap` controls the horizontal spacing between year blocks

## Example: Both vertical and  Horizontal whitespace

```{r}
plot_sankey_stroke(
  stroke_par_blacks,
  vertical_gap = 0.02,
  horizontal_gap = 0.05
)
```

## Example: No horizontal whitespace

```{r}
plot_sankey_stroke(
  stroke_par_blacks,
  vertical_gap = 0.02,
  horizontal_gap = 0
)
```

## Example: No vertical whitespace

```{r}
plot_sankey_stroke(
  stroke_par_blacks,
  vertical_gap = 0,
  horizontal_gap = 0.05
)
```

## Example: No whitespace

```{r}
plot_sankey_stroke(
  stroke_par_blacks,
  vertical_gap = 0,
  horizontal_gap = 0
)
```

# Summary

This package demonstrates how Sankey-style diagrams can be created using `ggplot2` with custom geometry. The plotting function also provides parameters that allow users to control the whitespace between visual elements.
